Individual Variation in Animal Movement and Effects on Seed Dispersal Distance

### 2021-02-01 paper/figures_tables.Rmd
Methods and figures up to methods are ready!

#### Figure 3.
Variation in Weibull and Generalized Pareto distributions determined by the value of the shape parameter. **A.** In the case of the Weibull distribution, the scale parameter is defined as $b=1$, and the shape parameters are defined as follows: solid line $a=\nu=0.4$ shows a heavy tail, the dashed line $a=\nu=2$ shows a thin tail, and $a=\nu=1$ describes an exponential tail with the dotted line. This follows the assumption that a shape parameter $\nu<1$ describes a fat tail, $\nu>1$ a thin tail, and $\nu=1$ a tail following an exponential distribution. *Check here these statements about the tail. Maybe parameterization has changed it, but I don't see the heavy tail of the solid line with $\beta > 1$* **B.** In the case of the Generalized Pareto distribution, a shape parameter $=0$ describes an exponential distribution function, shape $>0$ describes a heavy tail, and $<0$ a tail with Beta distribution function, bounded at an upper value as a function of the threshold, shape and scale parameters. In the second panel, the scale parameter is set to $\sigma = 1$ the solid line $\xi = 1$ shows a heavy tail, dashed line $\xi=-0.6$ shows a Beta tail, and $\xi = 0$ describes aan exponential decay with the dotted line.

>>>>

#### Figure 3.
Variation in Weibull and Generalized Pareto distributions determined by the value of the shape parameter. **A.** In the case of the Weibull distribution, the scale parameter is defined as $b=1$, and the shape parameters are defined as follows: solid line $a=\nu=0.4$ shows a heavy tail, the dashed line $a=\nu=2$ shows a thin tail, and $a=\nu=1$ describes an exponential tail with the dotted line. This follows the assumption that a shape parameter $\nu<1$ describes a fat tail, $\nu>1$ a thin tail, and $\nu=1$ a tail following an exponential distribution. **B.** In the case of the Generalized Pareto distribution, a shape parameter $\xi=0$ describes an exponential distribution function, shape $\xi>0$ describes a heavy tail, and $\xi<0$ a tail with Beta distribution function, bounded at an upper value as a function of the threshold and scale parameters. In the second panel, the scale parameter is set to $\sigma = 1$ the solid line $\xi = 1$ shows a heavy tail, dashed line $\xi=-0.6$ shows a Beta tail, and $\xi = 0$ describes an exponential decay with the dotted line.

### 2021-02-01 paper/paper.Rmd
Methods and figures up to methods are ready!

Where the shape parameter $a$ determines the tail of the distribution, and $b$ is the scale parameter. It is worth mentioning that the standard parameterization of the Weibull distribution in 'R' is different from the parameterization used in Morales and Carlo  [-@morales_effects_2006], where the shape parameter $a = \nu$ remains the same, but the scale parameter is defined as $\kappa = b^{-\nu}$, giving a density distribution function of $W(r)= \kappa\nu r^{\nu-1} exp[-{\kappa r}^\nu]$. 
<!-- Here, we report the scale parameter following the alternative parameterization, but the original values returned after fitting the Weibull distribution using R can be found in the Supplementary Table 1.  -->
Since the shape parameter remains the same, we can note that for $a=\nu=1$ the distribution shows a tail with exponential decay, with values of $\nu <1$ the tail shows fast-decay, and when $\nu>1$ we can see a fat-tailed distribution (Fig 3.) We visually assessed the fit of the distributions via qqplots and used the Kolmogorov-Smirnov statistic to assess goodness of fit.

Given our interest in the tail-end of the seed dispersal kernels and our focus on the occurrence of long-distance events, we decided to use an extreme value theory approach and fit a Generalized Pareto distribution to the dispersal distances above the 500m threshold established. Statistics of extremes are useful to understand the stochastic behavior of rare extreme events [@coles2001introduction] and have previously been used in ecological scenarios [@gaines1993largest; @katz2005statistics] and evolutionary genetics [@beisel2007testing; @joyce2008general]. More recently, statistics of extremes have been used to model extended dispersal kernels of seeds and pollen [@garcia_extended_2017] and in particular have been proposed to understand long-distance dispersal events [@rogers_total_2019]. In our particular case, we were interested to evaluate how an approach using statistics of extremes could help us fit the tail of the frugivore-generated seed dispersal kernels for each of the models we simulated. We used a peak over threshold (POT) approach, in which we filtered the dispersal distances above a given threshold and fit a Generalized Pareto (GP) distribution to those values. This differs from a block maxima (BM) approach, where the maximum value is selected for each sampling unit and a Generalized Extreme Value (GEV) distribution is used, therefore the choice in the approach is determined by the structure in sampling for a given study (further details can be found in [@coles2001introduction; @garcia_extended_2017]). Further details on these two approaches and applications to dispersal can be found [@garcia_extended_2017].

Our study focused on long-distance dispersal events defined as any dispersal event over 500m from the parent tree in order to compare our results to previous studies in this same ecological system [@holbrook_dispersal_2009; @holbrook_home_2011; @holbrook_using_2007]. However, when selecting a threshold for a POT approach, it is also possible to use diagnostic plots by fitting the data to a sequence of thresholds, with the goal of finding the lowest threshold that provides similar parameter values to any other higher threshold. We performed these analysis (Supplementary information), and given that parameter estimates were similar, we report our parameter estimates and expected probability of occurrence of long-distance events using our original 500m threshold. We fit Generalized Pareto distributions to the data via maximum likelihood using the package extRemes [@extRemes2] in R [Rcit].

>>>>

Where the shape parameter $a$ determines the tail of the distribution, and $b$ is the scale parameter. It is worth mentioning that the standard parameterization of the Weibull distribution in 'R' is different from the parameterization used in Morales and Carlo  [-@morales_effects_2006], where the shape parameter $a = \nu$ remains the same, but the scale parameter is defined as $\kappa = b^{-\nu}$, giving a density distribution function of $W(r)= \kappa\nu r^{\nu-1} exp[-{\kappa r}^\nu]$. 
<!-- Here, we report the scale parameter following the alternative parameterization, but the original values returned after fitting the Weibull distribution using R can be found in the Supplementary Table 1.  -->
Since the shape parameter remains the same, we can note that for $a=\nu=1$ the distribution shows a tail with exponential decay, with values of $\nu > 1$ the tail shows fast-decay or thin tail, and when $\nu<1$ we can see a fat-tailed distribution as shown in [@morales_effects_2006] (Fig 3a.) We visually assessed the fit of the distributions via qqplots and used the Kolmogorov-Smirnov statistic to assess goodness of fit.

Given our interest in the tail-end of the seed dispersal kernels and our focus on the occurrence of long-distance events, we decided to use an extreme value theory approach and fit a Generalized Pareto distribution to the dispersal distances above the 500m threshold established. Statistics of extremes are useful to understand the stochastic behavior of rare extreme events [@coles2001introduction] and have previously been used in ecological scenarios [@gaines1993largest; @katz2005statistics] and evolutionary genetics [@beisel2007testing; @joyce2008general]. More recently, statistics of extremes have been used to model extended dispersal kernels of seeds and pollen [@garcia_extended_2017] and in particular have been proposed to understand long-distance dispersal events [@rogers_total_2019]. In our particular case, we were interested to evaluate how an approach using statistics of extremes could help us fit the tail of the frugivore-generated seed dispersal kernels for each of the models we simulated. We used a peak over threshold (POT) approach, in which we filtered the dispersal distances above a given threshold and fit a Generalized Pareto (GP) distribution to those values. This differs from a block maxima (BM) approach, where the maximum value is selected for each sampling unit and a Generalized Extreme Value (GEV) distribution is used, therefore the choice in the approach is determined by the structure in sampling for a given study (further details can be found in [@coles2001introduction; @garcia_extended_2017]). The family of generalized Pareto distributions is [@coles2001introduction]:

$$H(y) = 1 - \left(1 + \frac{\xi y }{\tilde{\sigma}}\right)^{-1/\xi}$$

defined on $\{y : y>0$ and $1 + \xi y/\tilde{\sigma}\}$, where
$$\tilde{\sigma} = \sigma + \xi(x - \mu)$$
An approximation between the GEV and GP distributions occurs where the values over a threshold excess (location, $\mu$) have a corresponding approximate GP distribution [@coles2001introduction]. The parameters in the GP distribution of threshold excesses are determined by the values in the corresponding GEV distribution of a random variable $X$ (scale $\sigma$, location $\mu$). In both distributions, the shape parameter $\xi$ is dominant to determine the tail of the distribution. In the generalized Pareto distribution, a shape parameter $\xi<0$ the tail has an upper bound at $\mu - \tilde{\sigma}/\xi$ following a Beta distribution function;whereas a shape parameter $\xi>0$ follows a heavy tail with no upper limit. In the special case of $\xi = 0$, the GP distribution function approximates to an exponential distribution with a parameter of $1/\tilde{\sigma}$ (Figure 3b.) [@coles2001introduction].

Our study focused on long-distance dispersal events defined as any dispersal event over 500m from the parent tree in order to compare our results to previous studies in this same ecological system [@holbrook_dispersal_2009; @holbrook_home_2011; @holbrook_using_2007]. However, when selecting a threshold for a POT approach, it is also possible to use diagnostic plots by fitting the data to a sequence of thresholds, with the goal of finding the lowest threshold that provides similar parameter values to any other higher threshold. We performed these analysis (Supplementary information), and given that parameter estimates were similar, we report our parameter estimates and expected probability of occurrence of long-distance events using our original 500m threshold. We fit Generalized Pareto distributions to the data via maximum likelihood using the package extRemes [@extRemes2] in R [@Rcit].
